---
title: "PMLProject"
author: "BY"
date: "22/05/2020"
output:
  html_document: default
  'html_document: default': default
---

```{r setup, include=FALSE, results='hide'}
library(caret)
library(gbm)
library(randomForest)
library(rpart)
library(e1071)
library(rattle)
library(parallel)
library(doParallel)
```

## Project
It is now possible to collect a large amount of data about personal activity relatively inexpensively Using devices such as Jawbone Up, Nike FuelBand, and Fitbit. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

The goal of this project is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise. The participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The process of building the model, using cross validation, the expected out of sample error, the reasons for the selecting choices are explained in this report.

More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting  Dataset).
The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.


```{r download_data,include=FALSE}
traind<- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"), header = TRUE)
testd<- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"), header = TRUE)
dim(traind);dim(testd)
```
## Cleaning data
There are some variables in both datasets which include missing values. There are also some variables which looks have no effect on the outcome (user name, stamptime, window, etc.).Also, Some variables have no or very little variability in them and will likely not be good predictors. To identify those variables we used nearzerovar function. Then, We removed all of them.

```{r Preparing_data,include=FALSE}
traind1<- traind[,colSums(is.na(traind))==0]
traind2<- traind1[,-c(1:7)]
nsv<- nearZeroVar(traind2)
nsv
trainc<- traind2[,-nsv]
dim(trainc)


testd1<- testd[,colSums(is.na(testd))==0]
testd2<- testd1[,-c(1:7)]
testc<- testd2
dim(testc)
```

## Creating train and test data for prediction

```{r datapartition, echo=FALSE}
set.seed(3618)
inTrain<- createDataPartition(trainc$classe,
                              p=0.7, list=FALSE)
training<-trainc[inTrain,]
testing<- trainc[-inTrain,] 
dim(training);dim(testing)

```

## Building Model
To predict the outcome, we used the following methods:
1. Classification tree
2. Random forest, and
3. Gradient boosting.

We used cross-validation technique for evaluating the models.

## 1. Classification tree
Building a model with classification tree algorithm, we've found an accuracy of about 50% which is low. As a result, the model can't predict Classe with a proper accuracy using all other predictors. Generalization error or Out of sample error is 0.5 which is so high.

```{r Model_Tree, echo=FALSE}
modFit1<- train(classe ~ ., method="rpart",data=training)
print(modFit1$finalModel)
print(modFit1)
fancyRpartPlot(modFit1$finalModel)
 
# Validating the model: Accuracy
predict1<- predict(modFit1,newdata=testing)
conf1<- confusionMatrix(testing$classe,predict1)
conf1
```

## 2. Random forest
Building a model with random forest algorithm, we've found an accuracy of about 99.3% which is high. As a result, the model can predict Classe with a proper accuracy using all other predictors. Generalization error or Out of sample error is 0.01 which is quiet low. 

```{r Model_RandomForest, echo=FALSE}
fitControl<- trainControl(method="cv", number=5, verboseIter = FALSE)
modFit2<- train(classe ~ ., method="rf", data=training, trControl = fitControl)
print(modFit2$finalModel)
print(modFit2)
# Validating the model : Accuracy
predict2<- predict(modFit2,newdata=testing)
conf2<- confusionMatrix(testing$classe,predict2)
conf2
```
## 3. Gradient Boosting
Building a model with gradient boosting algorithm, we've found an accuracy of about 96.4% which is high. As a result, the model can predict Classe with a proper accuracy using all other predictors. Generalization error or Out of sample error is  0.04 which is quiet low. 

```{r Model_gbm, echo=FALSE}
fitControl<- trainControl(method="cv", number=5, verboseIter = FALSE)
modFit3<- train(classe ~ ., method="gbm", data=training, trControl = fitControl, verbose = FALSE)
print(modFit3$finalModel)
print(modFit3)
# Validating the model : Accuracy
predict3<- predict(modFit3,newdata=testing)
conf3<- confusionMatrix(testing$classe,predict3)
conf3
```
## Best model
We selected Random Forest model for prediction as it has the highest accuracy among all models.

We applied the model to test data then.

```{r model_valid, echo=FALSE}
predTest<- predict(modFit2, newdata = testc)
predTest
```
